\pkg{cv}: An \proglang{R} Package for Cross-Validating Regression Models

### 2024-07-09 scratch/notes_john/JSS-article/JSS-article-4.Rmd
finish first edit of JSS-article-4

The \pkg{cv} package is designed to be extensible in several directions. In order of increasing general complexity, we can add: (1) a cross-validation cost criterion; (2) a model class that's not directly accommodated by the \code{cv()} default method or by another directly inherited method; and (3) a new model-selection procedure suitable for use with the \code{"function"} method for \code{cv()}. In this section, we illustrate (1) and (2); more diverse and extensive examples may be found in the vignette on extending the \pkg{cv} package.

@@

<!-- We illustrate with an example adapted from a vignette in the \pkg{rsample} package, which uses the \code{attrition} data set from the \pkg{modeldata} package [@Kuhn:2024] (and which is by a wide margin the most downloaded from the RStudio CRAN mirror of the CV packages mentioned in the preceding paragraph).  -->

@@

The \pkg{caret} package supports generalized linear models fit by \code{glm()} and can straightforwardly perform LOO CV:
```{r rsample-example-using-caret, cache=TRUE}
library("caret", quietly=TRUE, warn.conflicts=FALSE)
# reports complement of Bayes-rule criterion:
train(x = attrition[, c("JobSatisfaction", "Gender",
                               "MonthlyIncome")],
             y = attrition$Attrition,
             method = "glm",
             trControl = trainControl(method="LOOCV"))
```

@@

Let's compare the computational efficiency of the various implementations, also showing the use of \code{method="Woodbury"} and \code{method="hatvalues"} for \code{cv()}, which, in this example, produce numerically identical results to the default \code{method="exact"}:
```{r compare-rsample-cv, cache=TRUE}
all.equal(cv.attrition$`CV crit`,
          cv(mod.attrition, k="loo", criterion=BayesRule,
             method="Woodbury")$`CV crit`,
          )
all.equal(cv.attrition$`CV crit`,
          cv(mod.attrition, k="loo", criterion=BayesRule,
             method="hatvalues")$`CV crit`
          )

print(microbenchmark::microbenchmark(
  caret=train(x = attrition[, c("JobSatisfaction", "Gender",
                                "MonthlyIncome")],
              y = attrition$Attrition,
              method = "glm",
              trControl = trainControl(method="LOOCV")),
  rsample={
    rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
    rs_obj$accuracy <- map_dbl(rs_obj$results,
                               function(x) mean(x$correct))
  },
  direct={
    n <- nrow(attrition)
    yhat <- numeric(n)
    for (i in 1:n){
      m <- update(mod.attrition, data=attrition[-i, ])
      yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
    }
    mean(((attrition$Attrition == "Yes") - round(yhat))^2)
  },
  cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
  cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
             method="Woodbury"),
  cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
                  method="hatvalues"),
  times=10,
  unit = "relative"
  ), signif=3)
```
Computing time for \pkg{rsample} and \pkg{caret} for this problem is similar to direct computation of LOO CV and to \code{cv()} with \code{method="exact"}, but almost two orders of magnitude slower than \code{cv()} with \code{method="Woodbury"}, and more than three orders of magnitude slower then \code{cv()} with \code{method="hatvalues"}.

>>>>

The \pkg{cv} package is designed to be extensible in several directions. In order of increasing general complexity, we can add: (1) a cross-validation cost criterion; (2) a model class that's not directly accommodated by the \code{cv()} default method or by another directly inherited method; and (3) a new model-selection procedure suitable for use with the \code{"function"} method for \code{cv()}. In this section, we illustrate (1) and (2); more diverse and extensive examples (including of (3)) may be found in the vignette on extending the \pkg{cv} package.

@@

@@

The \pkg{caret} package supports generalized linear models fit by \code{glm()} and can straightforwardly perform LOO CV:
```{r rsample-example-using-caret, cache=TRUE}
library("ggplot2", warn.conflicts=FALSE) # used by caret
library("caret", quietly=TRUE)
# reports complement of Bayes-rule criterion:
train(x = attrition[, c("JobSatisfaction", "Gender",
                               "MonthlyIncome")],
             y = attrition$Attrition,
             method = "glm",
             trControl = trainControl(method="LOOCV"))
```
where "accuracy" is the complement of the misclassification rate.

@@

Let's compare the computational efficiency of the various implementations, also showing the use of \code{method="Woodbury"} and \code{method="hatvalues"} for \code{cv()}, which, in this example, produce numerically identical results to the default \code{method="exact"}:[^times-10]

[^times-10]: As before, we ask \code{microbenchmark()} to report relative timings, and also as before, we set \code{times = 10} rather than the default \code{times = 100}, with the consequence that we fail to detect a statistically reliable difference between using the Woodbury matrix identity and using hatvalues for the \code{cv()} computations.

```{r compare-rsample-cv, cache=TRUE}
print(microbenchmark::microbenchmark(
  caret=train(x = attrition[, c("JobSatisfaction", "Gender",
                                "MonthlyIncome")],
              y = attrition$Attrition,
              method = "glm",
              trControl = trainControl(method="LOOCV")),
  rsample={
    rs_obj$results <- map(rs_obj$splits, holdout_results, mod_form)
    rs_obj$accuracy <- map_dbl(rs_obj$results,
                               function(x) mean(x$correct))
  },
  direct={
    n <- nrow(attrition)
    yhat <- numeric(n)
    for (i in 1:n){
      m <- update(mod.attrition, data=attrition[-i, ])
      yhat[i] <- predict(m, newdata=attrition[i, ], type="response")
    }
    mean(((attrition$Attrition == "Yes") - round(yhat))^2)
  },
  cv.exact=cv(mod.attrition, k="loo", criterion=BayesRule),
  cv.wood=cv(mod.attrition, k="loo", criterion=BayesRule,
             method="Woodbury"),
  cv.hatvalues=cv(mod.attrition, k="loo", criterion=BayesRule,
                  method="hatvalues"),
  times=10,
  unit = "relative"
  ), signif=3)
```
Computing time for \pkg{rsample} and \pkg{caret} for this problem is similar to direct computation of LOO CV and to \code{cv()} with \code{method="exact"}, but almost two orders of magnitude slower than \code{cv()} with \code{method="Woodbury"}, and more than three orders of magnitude slower then \code{cv()} with \code{method="hatvalues"}.
